---
title: "M-out-of-N Bootstrap"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

**Purpose**

* The purpose of this script is to geneate m-ou-of-n bootstrap standard error estimates for linear regression models of the mean pollutant concentration in a block based on the proportion of each race within that block, such that: 

$$pollutant\ conc = \beta_1 X_{race-ethnicity\ pct}$$

* We will take an m-ou-of-n bootstrap approach. The general approach here is: 




$SE = \frac{\hat\sigma}{\sqrt n}$

where $s$ is the standard deviation from the ____??? bootstrap sample, and ___??? $n$ is the bootstrap sample size


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=F, cache.comments = F, 
                      message = F, warning = F, tidy.opts=list(width.cutoff=60), tidy=TRUE, 
                      fig.height = 6, fig.width = 10
                      )  


# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(#kableExtra, 
               tidyverse,
               #boot #bootstrap fn
               broom
               )

set.seed(1)

#prevent scientific notation
options(scipen=999)

image_path <- file.path("..", "Manuscript", "Images")

#add legends to bottom of plots
#theme_update(legend.position="bottom")

```

# Functions

```{r}
# function returns dataframe with appropriate labels 

add_labels <- function(dt, var = "variable", race_var = "race", keep_original_var=TRUE) {
  
  dt <- dplyr::rename(dt, var=var,
                      race_var = race_var
  )
  
  if(keep_original_var==TRUE){
    dt <- mutate(dt,
                 var0 = var,
                 race_var0 = race_var
    )
    }
  
  dt <- dt %>%
      mutate(
        # pollutants 
        var = recode_factor(factor(var),
                            "co2_umol_mol" = "CO2 (ppm)",
                            "ma200_ir_bc1" = "BC (ng/m3)",
                            "pm2.5_ug_m3" = "PM2.5 (ug/m3)",
                            "no2" = "NO2 (ppb)",
                            "pnc_noscreen" = "PNC (pt/cm3)",
                            ),
         var = factor(var, levels = c("PNC (pt/cm3)",
                                      "BC (ng/m3)",
                                      "NO2 (ppb)",
                                      "PM2.5 (ug/m3)",
                                      "CO2 (ppm)")),
         
         #race-ethnicity groups
         race_var = recode_factor(factor(race_var),
                        "bk_p_race_asian" = "Asian",
                        "bk_p_race_black" = "Black",
                        "bk_p_ethn_hisp" = "Hispanic",
                        "bk_p_ethn_non_hisp_white" = "Non-Hispanic White"),
         race_var = factor(race_var, levels = c("Asian", "Black", "Hispanic", "Non-Hispanic White"))
        )
  
  names(dt)[names(dt) == "var"] <- var
  names(dt)[names(dt) == "race_var"] <- race_var
  
  return(dt)
  
  }
  

```


# Load Data 

```{r}
# load data 
## census data
census_file <- file.path("..", "..", "dr0342_census_block")
if(file.exists(paste0(census_file, ".rda"))) {
  census <- readRDS(paste0(census_file, ".rda"))
  } else {
    census <- read.csv(paste0(census_file, ".csv")) %>% select(-X)
    saveRDS(census, paste0(census_file, ".rda"))
    }

#pollutants of interest
unique_pollutants <- c("pnc_noscreen", "ma200_ir_bc1", "no2", "pm2.5_ug_m3") #unique(predictions$variable)

## model predictions for air pollution
predictions0 <- readRDS(file.path("..", "..", "Block Predictions", "20220207", "predictions.rda")) %>%
  # only keep predictions in the modeing area
  filter(in_monitoring_area ==TRUE,
         variable %in% unique_pollutants) %>%
  select(-c(in_monitoring_area, annual, msa)) %>%
  # there were some differences between the originally available 2000 census block IDs and the 2010 census block IDs used in this analysis. this addresses that.
  mutate(block_key = as.numeric(substr(native_id, 1, 15)))


race_variables <- c("bk_p_race_asian", "bk_p_race_black", "bk_p_ethn_hisp", "bk_p_ethn_non_hisp_white")


### length(unique(predictions0$block_key)) # 29,521
### length(unique(census$block_key)) # 195,574
#merge by "block_key"

predictions <- left_join(predictions0, census) %>%
  # some blocks have NAs because nobody lives there (e.g., national parks)
  drop_na()

predictions_long <- predictions %>%
  select(block_key, variable, prediction, all_of(race_variables)) %>%
  gather("race", "race_prop", race_variables) 
```


```{r, eval=F}
# TEMP
# map showing blocks w/ NAs
predictions %>%
  mutate(has_na = is.na(bk_p_race_black)) %>%
           ggplot(aes(x=longitude, y=latitude, col=has_na)) + 
           geom_point()
         
         
         
```


```{r}
# common variables

#unique blocks 
unique_blocks <- unique(predictions$block_key)
# total number of blocks #10k
n_blocks <- length(unique_blocks)
# blocks to sample in each replication 
m <- 2*sqrt(n_blocks) %>% round()
#m <- 1*sqrt(n_blocks) %>% round()
# number of bootstrap samples to take
replicate_n <- 500  

```



# Linear regression models

## air pollution ~ race 

```{r}
#original Lms 

original_race_lms <- lapply(group_split(predictions_long, variable, race), function(x){
  # x = group_split(predictions_long, variable, race)[[1]]
  
  lm1 <- lm(prediction~race_prop, data=x) %>%
    tidy(conf.int=T) %>%  
    mutate(
      variable = first(x$variable),
      race = first(x$race)
    )
  }) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(CI = "standard")

```

 

Approach for m-of-n bootstrap linear regression models:  

### --> SIZE of m? 

1. randomly sample m census blocks with replacement, where $m = 2\sqrt n$. Conduct 500 bootstrap sample replicates. 
2. run the appropriate pollution regression models with each bootstrap sample 

3. calculate the mean and 95% confidence interval for each race-ethnicity pollution model as the mean, 2.5th, and 97.5th quantile of the 500 bootstrapped coefficient estimates 



```{r}

bootstrap_lm <- function(dt = predictions_long, boot_sample_size=m) {
  
  block_sample <- sample(unique(dt$block_key), size = boot_sample_size, replace = T)
  temp_df <- filter(dt, block_key %in% block_sample)

  lms <- lapply(group_split(temp_df, variable, race), function(x){

    lm(prediction~race_prop, data= x) %>%
    tidy(conf.int=T) %>%
    mutate(
      variable = first(x$variable),
      race = first(x$race)
    )
  }) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") 

  return(lms)
  
}

 
# calculate the 95% CI
set.seed(1)
boot_race_lms0 <- replicate(replicate_n, 
          expr = bootstrap_lm(), 
          simplify = F) %>%
  bind_rows() 

boot_race_lms <- boot_race_lms0 %>%
  group_by(variable, race) %>%
  summarize(
    # estimate using the coefficient estimates
    conf.low = quantile(estimate, 0.025),
    conf.high = quantile(estimate, 0.975),
    
    estimate = mean(estimate),
    
    # pulling the SE approach
    #std.error = mean(std.error),
    # # this produces very similar results are below (estiamte +- t-score*se)
    # conf.low2 = mean(conf.low),
    # conf.high2 = mean(conf.high)
    ) %>%
  mutate(CI = "bootstrapped") 
   
#merge results 
all_lms <- original_race_lms %>%
  select(variable, race, estimate, #std.error, 
         conf.low, conf.high, CI) %>%
  bind_rows(boot_race_lms)  %>%
  add_labels(keep_original_var = F)


```


```{r}
pd <- position_dodge(0.4)

```

bootstrapped esimate and CIs

```{r}
all_lms %>%
  filter(CI == "bootstrapped") %>% 
  ggplot(., aes(x = estimate, y=race, #col=CI,
                col=variable,
                    xmin = conf.low, xmax = conf.high, height = 0,
                    )) +
  facet_grid(cols = vars(variable), rows = vars(race), scales="free", switch="y") +
  geom_vline(xintercept = 0, lty = 2, alpha=0.5) +
  geom_point(position = pd) +
  geom_errorbarh(position = pd) + 
  theme(legend.position = "none") +
  labs(y ="Race-Ethnicity (%)", x = "Coefficient Estimate"#, #col= "95% CI"
       ) +
  scale_y_discrete(labels=NULL, breaks=NULL)

ggsave(file.path(image_path, "race_boot_lms.png"),width = 8, height = 6)

```


Comparison plot of bootstrapped and standared 95% confidence intervals
 
```{r}

ggplot(all_lms, aes(x = estimate, y=race, col=CI,
                    xmin = conf.low, xmax = conf.high, height = 0,
                    )) +
  facet_grid(cols = vars(variable), rows = vars(race), scales="free", switch="y") +
  geom_vline(xintercept = 0, lty = 2, alpha=0.5) +
  geom_point(position = pd) +
  geom_errorbarh(position = pd) + 
  labs(y ="Race-Ethnicity (%)", x = "Coefficient Estimate", col= "95% CI") +
  scale_y_discrete(labels=NULL, breaks=NULL)

ggsave(file.path(image_path, "comparison.png"),width = 8, height = 6)

```



 











```{r}
knitr::knit_exit()
```



