---
title: "M-out-of-N Bootstrap"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

**Purpose**

* The purpose of this script is to geneate m-ou-of-n bootstrap standard error estimates for linear regression models of the mean pollutant concentration in a block based on the proportion of each race within that block, such that: 

$pollutant\ conc = \beta_1 X_{race-ethnicity\ pct}   $

* We will take an m-ou-of-n bootstrap approach. The general approach here is: 




$SE = \frac{\s}{\sqrt n}$

where $s$ is the standard deviation from the ____??? bootstrap sample, and ___??? $n$ is the bootstrap sample size


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=F, cache.comments = F, 
                      message = F, warning = F#, tidy.opts=list(width.cutoff=60), tidy=TRUE, fig.height = 10, fig.width = 10
                      )  


# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(#kableExtra, 
               tidyverse,
               #boot #bootstrap fn
               broom
               )

set.seed(1)

#prevent scientific notation
options(scipen=999)

```

# Load Data 

```{r}
# load data 
## census data
census_file <- file.path("..", "..", "dr0342_census_block")
if(file.exists(paste0(census_file, ".rda"))) {
  census <- readRDS(paste0(census_file, ".rda"))
  } else {
    census <- read.csv(paste0(census_file, ".csv")) %>% select(-X)
    saveRDS(census, paste0(census_file, ".rda"))
    }

#pollutants of interest
unique_pollutants <- c("pnc_noscreen", "ma200_ir_bc1", "no2", "pm2.5_ug_m3") #unique(predictions$variable)



## model predictions for air pollution
predictions0 <- readRDS(file.path("..", "..", "Block Predictions", "20220207", "predictions.rda")) %>%
  # only keep predictions in the modeing area
  filter(in_monitoring_area ==TRUE,
         variable %in% unique_pollutants
         ) %>%
  select(-c(in_monitoring_area, annual, msa)) %>%
  # kaya does this - don't need to do this, though?
  mutate(block_key = as.numeric(substr(native_id, 1, 15)))


## ??? these have different number of unique block_key values?? 
## merge by block_key ???

```

### --> ?? block_key is native_id? 
### --> why are there 10k block_keys but 15k native_id and location_id values? use native_id?

### --> why are there blocks w/ missing race data?? is the merging wrong?

```{r}
race_variables <- c("bk_p_race_asian", "bk_p_race_black", "bk_p_ethn_hisp", "bk_p_ethn_non_hisp_white")


### length(unique(predictions0$block_key)) # 29,521
### length(unique(census$block_key)) # 195,574
#merge by "block_key"

predictions <- left_join(predictions0, census) %>%
  
  ### ---> FIX THIS LATER. there shouldn't be any blocks w/ missing data (?)
  drop_na()

predictions_long <- predictions %>%
  select(block_key, variable, prediction, all_of(race_variables)) %>%
  gather("race", "race_prop", race_variables)

```


### --> souldnt the number of unique blcoks be 15k (not 10k)? 

```{r}
# common variables

#predictions %>% group_by(block_key) %>% mutate(n=n()) %>% filter(n>4) %>% View()
  
#unique blocks 
unique_blocks <- unique(predictions$block_key)
# total number of blocks #10k
n_blocks <- length(unique_blocks)
# blocks to sample in each replication 
m <- 2*sqrt(n_blocks) %>% round()
# number of bootstrap samples to take
replicate_n <- 100#500

```

# Linear regression models 

## air pollution ~ race 

### --> why are my estimates different than Kaya's? she needs winsorized means? 

```{r}
#original Lms 

original_race_lms <- lapply(group_split(predictions_long, variable, race), function(x){
  # x = group_split(predictions_long, variable, race)[[1]]
  
  lm1 <- lm(prediction~race_prop, data=x) %>%
    tidy(conf.int=T) %>%  
    mutate(
      variable = first(x$variable),
      race = first(x$race)
    )
  }) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") %>%
  mutate(SE = "standard")

```

 

Bootstrap sampling census blocks, building regression models with each set; repeat 500 tiems

### --> SIZE of m? 
### --> sample w/o replacement?     
1. randomly sample m census blocks ??___WITHOUT__ replacement, where $m = 2\sqrt n$, 500 replicates 

### --> calculate CI as: boot_mean +- boot_se*t_statistic ? or the estimated CI from each simulation? same results?
### --> use t_score <- qt(p = 0.05/2, df = m-1, lower.tail = F) instead of 1.96?

### --> ?   
2. run regression models with each bootstrap sample 

2. calculate the mean SE from the bootstrapped sample results 

### --> use the bootstrapped mean coefficient estimate here as well (vs estimate from traditional regression)?   
2. calculate the bootstrapped CI:

$$CI = \bar X \pm t_{\alpha/2, N-1}SE_{\bar X} $$


```{r}
#dt = predictions_long
# boot_sample_size=m

bootstrap_lm <- function(dt = predictions_long, boot_sample_size=m) {
  
  block_sample <- sample(unique(dt$block_key), size = boot_sample_size, replace = F)
  temp_df <- filter(dt, block_key %in% block_sample)

  lms <- lapply(group_split(temp_df, variable, race), function(x){

    lm(prediction~race_prop, data= x) %>%
    tidy() %>%
    mutate(
      variable = first(x$variable),
      race = first(x$race)
    )
  }) %>%
  bind_rows() %>%
  filter(term != "(Intercept)") 

  return(lms)
  
}

 
# calculate the 95% CI

## tutorial: https://bookdown.org/logan_kelly/r_practice/p09.html 
  
#### --> ????
t_score <- qt(p = 0.05/2, df = m-1, lower.tail = F)

# sample a subset of blocks 500 times, run a model each time for each pollutant & race; calculate the mean SE & confidence interval

boot_race_lms <- replicate(replicate_n, 
          expr = bootstrap_lm(), 
          simplify = F) %>%
  bind_rows() %>%
  group_by(variable, race) %>%
  summarize(
    #need this??
    estimate = mean(estimate),
    
    std.error = mean(std.error)
    ) %>%
  mutate(
    conf.low = estimate - t_score*std.error,
    conf.high = estimate + t_score*std.error
  ) %>%
  mutate(SE = "boot")


```

merge results 

```{r}
all_lms <- original_race_lms %>%
  select(variable, race, estimate, std.error, conf.low, conf.high, SE) %>%
  bind_rows(boot_race_lms)  

```


plot
 
```{r}
# d <- tidy(mod, conf.int = TRUE)
# 
pd <- position_dodge(0.4)

ggplot(all_lms, aes(x = estimate, 
                    y=race,
                    #y = SE, 
                    col=SE,
                    xmin = conf.low, xmax = conf.high, height = 0,
                    )) +
  facet_grid(cols = vars(variable), scales="free") +
  geom_vline(xintercept = 0, lty = 2, alpha=0.5) +
  geom_point(position = pd) +
  geom_errorbarh(position = pd)

```



 


```{r}
knitr::knit_exit()
```











 




 



```{r, eval=F}

###--> ??? how to calculat SE for each group race?? 


bootstrap_se <- lapply(group_split(predictions_long, variable, race) , function (x) {
  # x = group_split(predictions_long, variable, race)[[1]]
  
  # returns a matrix where rows are # blocks sampled, columns are the replicates
  bs_samples <- replicate(replicate_n, sample(x =x$race_prop, size = m, replace = T))
  
  # ??? calculate the SE of each replicate, then take the mean of all the replicates
  ## length(y) = m here
  bs_se <- apply(bs_samples, 2, function(y) { sqrt(var(y)/length(y))}) %>%
    mean()
  
  # overall mean proportion 
  bs_mean <- mean(bs_samples)
  
  
  
  
  # calculate the 95% CI
  ## tutorial: https://bookdown.org/logan_kelly/r_practice/p09.html 
  
  #### --> ????
  ## 1.97 , not 1.96 ??
  t_score <- qt(p = 0.05/2, df = m-1, lower.tail = F)
  margin_error <- t_score*bs_se
  
  #CI
  ci_lower_bound = bs_mean - margin_error
  ci_upper_bound = bs_mean + margin_error
  
  #return the results as a data frame
  data.frame(
    variable = first(x$variable),
    race = first(x$race),
    bs_mean = bs_mean,
    bs_se = bs_se,
    ci_lower_bound = ci_lower_bound,
    ci_upper_bound = ci_upper_bound
  )
  }) %>%
  bind_rows()

bootstrap_se

```


 



 


 